---
title: "Subreddit Diet Pipeline — 1. Import"
subtitle: "Import the dedicated subreddits — inflammatory myopathies (myositis + dermatomyositis) vs muscular dystrophies (muscular_dystrophy + FSHD)"
author: "Peter Kerkhof"
date: last-modified
format:
  html: default
  pdf: default
  docx: default
editor: visual
execute:
  echo: false
  warning: false
editor_options:
  chunk_output_type: inline
always_allow_html: yes
---

## Overview

Imports the Arctic Shift JSONL archives for the dedicated disease subreddits, grouped into two disease groups — inflammatory myopathies (r/myositis, r/dermatomyositis) vs muscular dystrophies (r/muscular_dystrophy, r/FSHD) — and merges them. The full history of each subreddit is kept.

**A note on coverage, to be carried into the paper's limitations.** The four subreddits were not all active over the same span, and this is real community history, not a gap in the data — we verified it against a second, independent copy of Reddit (the Pushshift archive) and against the live site.

- **r/dermatomyositis** is locked. Reddit marks it "restricted" and its newest submission is from 10 July 2020; after that it only collects comments on old threads. So its posts are concentrated in 2019–2020.
- **r/Myositis** was founded in 2020 and stayed nearly empty until early 2024, then grew quickly. Most of its posts are from 2024 onwards.
- **r/muscular_dystrophy** and **r/FSHD** run steadily across the whole period.

The consequence is that the two sides of the comparison are not observed evenly over time: the muscular-dystrophy side spans the full window smoothly, while the myositis side is front-loaded (dermatomyositis) and back-loaded (r/Myositis). A same-window analysis is possible but leaves too few myositis posts to be worth it, so we keep the full history for statistical power and state this temporal imbalance plainly as a limitation. Any difference between the two groups could in part reflect *when* each community was active, not only *which* disease it represents.

## 1. Setup

```{r}
#| message: false
#| label: setup-environment

rm(list = ls()[!sapply(ls(), function(x) is.function(get(x)))])
cat("\014")
options(scipen = 999)

# Keep non-ASCII characters (>=, em-dash) intact in figures and printed output.
# Running from a bare terminal can start R in a non-UTF-8 locale, which silently
# turns them into "<U+2265>" in plot labels and console output. RStudio already
# sets a UTF-8 locale, so this is only a safety net for terminal/CI runs.
if (!isTRUE(l10n_info()[["UTF-8"]])) {
  invisible(suppressWarnings(Sys.setlocale("LC_CTYPE", "en_US.UTF-8")))
}

library(tidyverse)
library(jsonlite)
library(here)
library(arrow)
# Two notes on packages:
# 1. Don't also load the 'feather' package. Both it and 'arrow' provide functions with
#    the same names (read_feather / write_feather), so whichever loads last silently
#    wins — meaning the script's behaviour would depend on the order of these lines.
#    'arrow' is the modern replacement, so we use it on its own.
# 2. 'anonymizer' is used once (section 6) and is called directly as
#    anonymizer::anonymize(), so it doesn't need loading here.

source(here("config.R"))

# Helper: percentage table (normally in .Rprofile)
if (!exists("perc_table")) {
  perc_table <- function(df, column_name) {
    df |>
      count(!!rlang::sym(column_name), sort = TRUE, name = "count") |>
      mutate(
        relative_freq_pct = (count / sum(count)) * 100,
        cumulative_freq_pct = cumsum(relative_freq_pct),
        across(ends_with("_pct"), ~ round(.x, 1))
      )
  }
}
```

## 2. Import subreddit JSONL archives

```{r}
#| message: false
#| label: import-subreddit-archives

# Columns to retain from the JSONL archives (subset present across files)
keep_cols <- c(
  "subreddit", "created_utc", "author", "title", "selftext", "body",
  "link_id", "parent_id", "id", "retrieved_on", "score", "url"
)

folder_path_myo <- file.path(REDDIT_DATA_PATH, "subreddits/data/myositis")
folder_path_mud <- file.path(REDDIT_DATA_PATH, "subreddits/data/muscular_dystrophy")

files <- c(
  list.files(folder_path_myo, pattern = "\\.jsonl$", full.names = TRUE,
             recursive = TRUE, ignore.case = TRUE),
  list.files(folder_path_mud, pattern = "\\.jsonl$", full.names = TRUE,
             recursive = TRUE, ignore.case = TRUE)
)

read_jsonl <- function(file_path) {
  con <- file(file_path, open = "r")
  tryCatch(
    stream_in(con, verbose = FALSE),
    error = function(e) {
      message(sprintf("Error reading %s: %s", file_path, e$message))
      NULL
    },
    finally = if (isOpen(con)) close(con)
  )
}

# Clean and standardize columns inside map() before combining files.
# This prevents two common issues:
# 1. Inconsistent data types: Some files have extra columns or different formats 
#    (e.g., lists vs. empty NAs) that cause errors when combined. Limiting to 
#    'keep_cols' and forcing all fields to text (character) standardizes them first.
# 2. Memory crashes: Processing files one-by-one keeps only the small, cleaned 
#    data in memory. Reading all massive raw files first would crash R.

df_raw <- files |>
  map(\(f) {
    d <- read_jsonl(f)
    if (is.null(d) || nrow(d) == 0) return(NULL)
    d |>
      select(any_of(keep_cols)) |>
      mutate(across(everything(), as.character))
  }) |>
  list_rbind()

cat("Raw posts read from JSONL:", scales::comma(nrow(df_raw)), "\n")

# Throw out anything we've read twice.
# We read whole folders, so if the same archive file sits in two places — easily done
# when you re-download one and drop the copy somewhere else — every post in it would
# otherwise be counted twice, and the subreddit would silently look twice as busy as
# it is. Every Reddit post has its own id, so a repeated id means a repeated post.
# We match on the id AND the timestamp: ids are only guaranteed unique within their
# own kind, so in principle a comment and a submission could carry the same one, and
# they would never share a posting time.
n_before <- nrow(df_raw)
df_raw <- df_raw |> distinct(id, created_utc, .keep_all = TRUE)
if (n_before > nrow(df_raw)) {
  cat("Dropped", scales::comma(n_before - nrow(df_raw)),
      "duplicate posts (the same archive file read more than once)\n")
}

cat("Unique posts:", scales::comma(nrow(df_raw)), "\n\n")
perc_table(df_raw, "subreddit")
```

## 3. Group the four dedicated subreddits

```{r}
#| message: false
#| label: select-subreddits

# Constructs a Myositis column = general myositis + the
# inflammatory-myopathy subtype dermatomyositis; and an MD column = general
# muscular dystrophy + FSHD (a specific muscular dystrophy).

myositis_subs <- c("Myositis", "dermatomyositis")
md_subs       <- c("MuscularDystrophy", "FSHD")
target_subs   <- c(myositis_subs, md_subs)

df <- df_raw |>
  filter(subreddit %in% target_subs) |>
  mutate(disease_type = if_else(subreddit %in% md_subs,
                                "Muscular Dystrophy", "Myositis"))

cat("Posts in the four dedicated subreddits:", scales::comma(nrow(df)),
    "| dropped by the allow-list:", scales::comma(nrow(df_raw) - nrow(df)), "\n\n")
print(df |> count(subreddit, disease_type, sort = TRUE))
```

## 4. Temporal parsing and filter

```{r}
#| message: false
#| label: temporal

# Keep the full history. Everything BEFORE this date is kept, so the corpus ends on
# the day before (14 July 2026).

cut_off_date <- "2026-07-15"

df <- df |>
  mutate(
    created_utc  = as.POSIXct(as.numeric(created_utc),  origin = "1970-01-01", tz = "UTC"),
    retrieved_on = as.POSIXct(as.numeric(retrieved_on), origin = "1970-01-01", tz = "UTC"),
    date     = date(created_utc),
    year     = year(created_utc),
    month    = month(created_utc),
    month_yr = format_ISO8601(created_utc, precision = "ym"),
    week_number = isoweek(date)
  ) |>
  arrange(created_utc) |>
  filter(date < as.Date(cut_off_date))

cat("Date range:", format(min(df$date)), "to", format(max(df$date)), "\n\n")

# Coverage by subreddit and year. This is the table behind the limitation noted at
# the top: dermatomyositis stops in 2020 (locked), r/Myositis only fills in from 2024,
# while the two dystrophy subreddits run throughout. Print it so the imbalance is
# visible in the rendered document, not just asserted in prose.
cat("Posts per subreddit per year (documents the temporal imbalance):\n")
print(df |> count(subreddit, year) |>
        pivot_wider(names_from = subreddit, values_from = n, values_fill = 0L) |>
        arrange(year), n = 30)
```

## 5. Text preprocessing

```{r}
#| message: false
#| label: preprocess-text

# Merge title + selftext + body into one lowercase text field with proper punctuation handling.

df <- df |>
  mutate(
    selftext = str_trim(selftext),
    title    = str_trim(title),
    body     = str_trim(body),
    title_punctuated = str_detect(title, "[\\.\\?\\!\\:]$"),
    title_clean = case_when(
      is.na(title) | title == "" ~ "",
      title_punctuated ~ paste0(title, " "),
      TRUE ~ paste0(title, ". ")
    ),
    text = paste0(title_clean, coalesce(selftext, ""), coalesce(body, "")),
    text = tolower(text),
    text = str_replace_all(text, "[-_]", " ")
  ) |>
  select(-title_punctuated, -title_clean) |>
  select(disease_type, subreddit, created_utc, text, everything())
```

## 6. Post URLs, anonymization, text metrics

```{r}
#| message: false
#| label: urls-metrics-anon

df <- df |>
  mutate(
    post_url = paste0("https://www.reddit.com/comments/", id),
    n_words  = str_count(text, "\\w+")
  )

# Anonymize author names using CRC32 hashing.
# # How it works:
# 1. Stable & Reproducible: Using a fixed seed (.seed = 0) ensures that the same 
#    username always gets the exact same anonymous ID every time this script runs.
# 2. Deleted Users: Users listed as "[deleted]" will also get a random-looking hash, 
#    but we manually change that back to a readable label later in the script.

df$author_ano <- anonymizer::anonymize(
  df$author, .algo = "crc32", .seed = 0)
  
df <- df |>
  mutate(author_ano = str_replace(author_ano, "3c0a4086", "deleted"))

```

## 7. Classify post types as submissions and comments

```{r}
#| message: false
#| label: classify-post-types

# Work out where each post sits in a conversation. Reddit threads have three levels:
#   submission          - someone starts a thread ("I was just diagnosed with DM...")
#   comment             - a reply to that thread starter
#   comment_to_comment  - a reply to someone else's reply
#
# How to tell them apart: every post records its "parent" (what it replies to), and
# Reddit labels IDs with a two-letter prefix — t3_ means the parent is a thread
# starter, t1_ means the parent is another comment. So:
#   no parent at all -> it IS the thread starter (a submission)
#   parent is t3_    -> a comment
#   parent is t1_    -> a reply to a reply
#
# Note that the archives sometimes store the parent ID without its prefix. When
# that happens we work it out by comparing with link_id, which always holds the ID of
# the thread the post belongs to. If the bare parent ID matches the thread ID, the
# post is replying to the thread starter (t3_); otherwise it is replying to a
# comment (t1_).

df <- df |>
  mutate(
    parent_id_chr = str_trim(coalesce(as.character(parent_id), "")),
    link_id_chr   = str_trim(coalesce(as.character(link_id), "")),
    parent_fullname = case_when(
      parent_id_chr == "" ~ NA_character_,
      str_starts(parent_id_chr, "t1_") | str_starts(parent_id_chr, "t3_") ~ parent_id_chr,
      parent_id_chr == str_remove(link_id_chr, "^t3_") ~ str_c("t3_", parent_id_chr),
      TRUE ~ str_c("t1_", parent_id_chr)
    ),
    post_type = case_when(
      is.na(parent_fullname) | parent_fullname == "" ~ "submission",
      str_starts(parent_fullname, "t3_") ~ "comment",
      str_starts(parent_fullname, "t1_") ~ "comment_to_comment",
      TRUE ~ "unknown"
    )
  ) |>
  select(-parent_id_chr, -link_id_chr)

print(df |> count(post_type, sort = TRUE) |>
        mutate(pct = scales::percent(n / sum(n), accuracy = 0.1)))
```

## 8. Save output

```{r}
#| message: false
#| label: save-output

output_path <- file.path(DATA_PATH, "data/myo_dys_subreddits_2005_jul2026.feather")
write_feather(df, output_path)
message(sprintf("Saved %s rows to: %s", scales::comma(nrow(df)), output_path))
```

## Output summary

- **Total posts**: `r scales::comma(nrow(df))`
- **Date range**: `r min(df$date)` to `r max(df$date)`
- **Subreddits**: r/myositis, r/dermatomyositis (Myositis group) vs r/muscular_dystrophy, r/FSHD (MD group)
- **Output**: DATA_PATH/data/`r basename(output_path)`
